---
title: "A5 - DD trends"
author: "Andrew Dickinson"
date: "Winter 2021"
#date: "<br>`r format(Sys.time(), '%d %B %Y')`"
header-includes:
  - \usepackage{mathtools}
  - \DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
  - \usepackage{amssymb}
output: 
  html_document:
    toc: false
    toc_depth: 3  
    number_sections: false
    theme: flatly
    highlight: tango  
    toc_float:
      collapsed: true
      smooth_scroll: true 
---

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(tidyverse, ggplot)
# Define pink color
red_pink <- "#e64173"
turquoise <- "#20B2AA"
orange <- "#FFA500"
red <- "#fb6107"
blue <- "#3b3b9a"
green <- "#8bb174"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
slate <- "#314f4f"
rmd_green <- "#73d7ad"
rmd_pink <- "#d7739d"
# Dark slate grey: #314f4f

# A blank theme for ggplot
theme_empty <- theme_bw() + theme(
  line = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  plot.margin = structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_A4 <- theme_bw() + theme(
  line = element_blank(),
  panel.grid = element_blank(),
  rect = element_blank(),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10),
  axis.ticks = element_blank(),
  axis.title.x = element_text(angle = 0, vjust = 0.5),
  axis.title.y = element_text(angle = 90, vjust = 0.5),
  legend.position = "bottom",
  plot.title = element_text(size = 10, face = "bold"),
  axis.line = element_line(color="black", size = .5)
)
```

> Due date: 16 February 2021

# {.tabset .tabset-fade .tabset-pills}


---

**Part 1.** Demonstrate that "_parallel trends in the absence of treatment_" is not sufficient to justify the difference-in-differences estimator&mdash;that, in order to identify a treatment-induced change in some outcome with a DD strategy, we must assume that outcomes in treated and control units are parallel _throughout the period of analysis_. Since we can evidence common pre-treatment trends, it is, if anything, in the _presence_ of treatment where the assumption is most active, in a way. 

*Q:* What is the DD model prepared to well identify? (_Level differences._) 

*Q:* Where do changes in slope (say, subsequent to treatment) show up then? (_They end up in the DD estimate. Do we wnat it to? Is there theory suggesting that we should anticipate only a level shift, or a change in both level and slope?_)

<br>

---

**Part 2.** Produce an event study figure on the same DGP.

<br>

---

## DD sim

#### Function

```{r}
dd_iter = function(iter = 1, n_ind = 10, n_period = 11, event = 6) {
  n_obs = n_period*n_ind
    dat_iter <- tibble(
    iter = iter,
    id = rep(1:n_ind, len = n_obs),
    time = rep(1:n_period, each = n_obs / n_period),
    post = ifelse(time >= event, 1, 0),
    treated = rep(0:1, length = n_obs),
    treat = treated*post,
    group = ifelse(treated == 1 & time >= event, 1,
                    ifelse(treated == 1 & time < event, 2,
                           ifelse(treated == 0 & time >= event, 3, 4))
                    ),
    y = 2 + 2*(treated == 1) + (0.2*time) + 1.5*(treat == 1) + rnorm(n_obs),
    z = 2 + 2*(treated == 1) + (0.2*time) + 1.5*(treat == 1)*(time / event) + rnorm(n_obs)
    )
    return(dat_iter)
}
```

#### Simulation

```{r}
sim_list <- map(1:500, dd_iter, n_ind = 10)
sim_df <- bind_rows(sim_list) %>% mutate(iter_group = iter + (group / 100))
```

#### Visual

```{r}
event = 6

ggplot(data = sim_df %>% filter(iter == 1), aes(x = time, y = y, colour = factor(treated))) +
  geom_point(alpha = 1) +
  ylim(ymin = -1, ymax = 10) +
  geom_vline(xintercept = event, linetype = "dashed") +
  geom_smooth(aes(group = group), method = "lm", se = F, color = "black") +
  scale_x_continuous(breaks = seq(0, 11, 1))
```


```{r}
ggplot(data = sim_df, aes(x = time, y = y, colour = factor(treated))) +
  geom_point(alpha = 0.01) +
  geom_vline(xintercept = 5, linetype = "dashed") +
  ylim(ymin = -1, ymax = 10) +
  stat_smooth(geom='line', aes(group = iter_group), method = "lm", alpha=0.1, se=FALSE, color = "grey50") +
  geom_smooth(aes(group = group), method = "lm", se = F, color = "black") +
  scale_x_continuous(breaks = seq(0, 11, 1))
```




```{r}
mean_z = sim_df %>% group_by(group) %>% summarize(mean_z = mean(z))
mean_y = sim_df %>% filter(group == 1) %>% summarize(mean_y = mean(y))

ggplot(data = sim_df, aes(x = time, y = z, colour = factor(treated))) +
  geom_point(alpha = 0.01) +
  ylim(ymin = -1, ymax = 10) +
  geom_vline(xintercept = 5, linetype = "dashed") +
  stat_smooth(geom='line', aes(group = iter_group), method = "lm", alpha=0.1, se=FALSE, color = "grey50") +
  geom_smooth(aes(group = group), method = "lm", se = F, size = 0.5, color = "black") +
  geom_smooth(data = sim_df %>% filter(group == 1), aes(y = y), method = "lm", se = F, color = "red", size = 0.5, linetype = "dashed") +
  geom_point(aes(x = 8.5, y = as.numeric(mean_y[1,1])), color = "black") +
  geom_point(aes(x = 8.5, y = as.numeric(mean_z[1,2])), color = purple) +
  geom_point(aes(x = 3, y = as.numeric(mean_z[2,2])), color = red) +
  geom_point(aes(x = 8.5, y = as.numeric(mean_z[3,2])), color = orange) +
  geom_point(aes(x = 3, y = as.numeric(mean_z[4,2])), color = green) +
  scale_x_continuous(breaks = seq(0, 11, 1))

```


## Event study


#### Simulation

```{r}
n_ind = 10
n_period = 10
n_obs = n_period*n_ind
event = 5

dat_iter = tibble(
  id = rep(1:n_ind, len = n_obs),
  time = rep(1:n_period, each = n_obs / n_period),
  post = ifelse(time >= event, 1, 0),
  treat = rep(0:1, length = n_obs),
  group = ifelse(treat == 1 & time >= event, 1,
                 ifelse(treat == 1 & time < event, 2,
                        ifelse(treat == 0 & time >= event, 3, 4))
  ),
  y = 2 + 2*(treat == 1) + (0.2*time) + 1.5*(treat == 1)*(post == 1) + rnorm(n_obs),
  z = 2 + 2*(treat == 1) + (0.2*time) + 1.5*(treat == 1)*(post == 1)*(time / event) + rnorm(n_obs)
) %>% panel(panel.id = ~id + time) %>% 
  arrange(id, time)
## run each model; treatment at i and c

m1 = feols(data = dat_iter, y ~ i(treat, time, 5) | id + time)
fixest::coefplot(m1)

m2 = feols(data = dat_iter, z ~ i(treat, time, 5) | id + time)
fixest::coefplot(m2)
```


#### Visual


