---
title: "A1 - OVB simulation"
author: "Andrew Dickinson"
date: "Winter 2020"
#date: "<br>`r format(Sys.time(), '%d %B %Y')`"
header-includes:
  - \usepackage{mathtools}
  - \DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
  - \usepackage{amssymb}
output: 
  html_document:
    toc: false
    toc_depth: 3  
    number_sections: false
    theme: flatly
    highlight: tango  
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(broom, latex2exp, leaflet, ggplot2, ggthemes, viridis, dplyr, magrittr, knitr, parallel, rddtools, readxl, emoGG, dslabs, gapminder, extrafont, Ecdat, wooldridge, tidyverse, janitor, kableExtra, gridExtra, estimatr, furrr, here, broom)
# Define pink color
red_pink <- "#e64173"
turquoise <- "#20B2AA"
orange <- "#FFA500"
red <- "#fb6107"
blue <- "#3b3b9a"
green <- "#8bb174"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
slate <- "#314f4f"
# Dark slate grey: #314f4f
# Notes directory
dir_slides <- "~/Dropbox/Courses/"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 4,
  fig.width = 6,
  # dpi = 300,
  # cache = T,
  warning = F,
  message = F
)
# A blank theme for ggplot
theme_empty <- theme_bw() + theme(
  line = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  plot.margin = structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_simple <- theme_bw() + theme(
  line = element_blank(),
  panel.grid = element_blank(),
  rect = element_blank(),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  axis.title.x = element_text(angle = 0, vjust = 0.5),
  axis.title.y = element_text(angle = 90, vjust = 0.5),
  legend.position = "none",
  axis.line = element_line(color="black", size = .5)
)
```

> Due date: 12 January 2020
>
> Please have your simulation ready to share in class on Tuesday.

__When people make the claim that _correlation does not imply causation_, they usually mean that the existence of some correlation between $y$ and $x_1$ does not imply that variation in $x_1$ _causes_ variation in $y$.__

- That is, they tend to be acknowledging that you can have correlation without causation

**Assignment: Propose and simulate a data-generating process in which (**i**) causation runs from $x_1$ to $y$ but at the same time (**ii**) the correlation of $y$ and $x_1$ is zero. Write it in an Rmd file and make the argument visually.**

- *Yes, at the end of this you should have simulated something (good) and demonstrated that the lack of correlation does not imply lack of causation (which is like a party trick).*

  
# {.tabset .tabset-fade .tabset-pills}

## Sim I

I want to propose and simulate a data generating process in which:
  - causation from $x_1$ to $y$
  - the correlation of $y$ and $x_1$ is zero

The first simulated DGP is a very very simple omitted variable example. I define my DGP as the following:

\begin{align*}
x_1 &= 1 + exp(0.01 * w) + e \\
x_2 &= 1 + 2.5x_1 + f \\
y &= 5x_1 - 2x_2 + u \\
\end{align*}
where
\begin{align*}
w &\sim N(5, 1), \\
u &\sim U(-5, 5), \\
e,f &\sim N(0,1) \\
\end{align*}

I simulate the DGP with the following function:

```{r sim1, cache=T}
# Set up function for a single iteration
sim1_iter = function(n = 1000) {
  # Generate data
  dgp_it = tibble(
    u = runif(n, -5, 5),
    e = rnorm(n, mean = 0, sd = 1),
    f = rnorm(n, mean = 0, sd = 1),
    w = (rnorm(n, mean = 5, sd = 1))^2,
    x1 = 1 + exp(0.01*w) + e,
    x2 = 1 + 2.5*x1 + f,
    y = 5*x1 - 2*x2 + u
  )
  # Collect estimates from regressions with and without OV
  bind_rows(
    lm(y ~ x1, data = dgp_it) %>% broom::tidy() %>% filter(term == "x1"),
    lm(y ~ x1 + x2, data = dgp_it) %>% broom::tidy() %>% filter(term == "x1"),
  ) %>% mutate(model = c("x2 is unobserved", "x2 is observed"))
}
```

Then simulating the function `sim1_iter()` 500 times with the following process:

```{r sim1_para, cache = T}
# Set up the parallelization
plan(multiprocess, workers = 8, .progress = T)
invisible(future_options(seed = 1234))
# Run the simulation
sim1_df = future_map_dfr(rep(15, 500), sim1_iter)
```

And I plot the resulting dataframe with `ggplot2`

```{r, cache = T}
# Plot simulation density plots
ggplot(data = sim1_df, aes(x = estimate, fill = model)) +
  geom_density(color = NA, alpha = 0.75) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 5, linetype = "dashed") +
  geom_vline(xintercept = 0, colour = "red") +
  labs(x = "Estimate", y = "Density") +
  scale_fill_viridis_d("Controls", option = "magma", end = 0.9) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```


## Sim II - Potential outcome framework
  
In my second simulation I write a DGP in a potential outcomes framework with a binary treatment variable $D_i$ and an outcome variable $Y_i$

I define my DGP as follows:

\begin{align*}
Y_{0i} &= X_i + u_i \\
Y_{1i} &= Y_{0i} + W_i + v_i \\
D_i &= \mathbb{I}(X_i + \varepsilon_i < 8) \\
Y_i &= Y_{0i} + D_i \tau_i \\
\end{align*}

where

\begin{align*}
u_i &\sim U(-5, 5) \\
v_i &\sim U(-3, 3) \\
\varepsilon_i &\sim U(-1, 1) \\
X_i &\sim N(10, 3) \\
W_i &\sim N(5, 1) \\

\end{align*}
  
```{r, sim2_iter, cache = T}
# Set up function for a single iteration
sim2_iter = function(n = 1000) {
  # Generate data
  dgp_it = tibble(
    u = runif(n, -5, 5),
    v = runif(n, -3, 3),
    e = runif(n, -1, 1),
    x = rnorm(n, mean = 10, sd = 3),
    w = rnorm(n, mean = 5, sd = 1),
    y0 = x + u,
    y1 = y0 + w + v,
    t = y1 - y0,
    d = (x + e < 8) %>% as.numeric(),
    y = y0 + d * t
  )
  # Collect estimates from regressions with and without OV
  bind_rows(
    lm(y ~ d, data = dgp_it) %>% broom::tidy() %>% filter(term == "d"),
    lm(y ~ d + x, data = dgp_it) %>% broom::tidy() %>% filter(term == "d"),
  ) %>% mutate(model = c("x is unobserved", "x is observed"))
}
```


```{r, sim2 para, echo = F, cache = T}
# Set up the parallelization
plan(multiprocess, workers = 8, .progress = T)
invisible(future_options(seed = 1234))
# Run the simulation
sim2_df = future_map_dfr(rep(15, 500), sim2_iter)
```


```{r sim2 plot, cache = T}
# Plot simulation density plots
ggplot(data = sim2_df, aes(x = estimate, fill = model)) +
  geom_density(color = NA, alpha = 0.75) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 5, linetype = "dashed") +
  geom_vline(xintercept = 0, colour = "red") +
  labs(x = "Estimate", y = "Density") +
  scale_fill_viridis_d("Controls", option = "magma", end = 0.9) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```



