---
title: "Micro-metrics"
author: "Andrew Dickinson"
date: "<br>`r format(Sys.time(), '%d %B %Y')`"
header-includes:
  - \usepackage{mathtools}
  - \DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
  - \usepackage{amssymb}
output: 
  html_document:
    toc: false
    toc_depth: 3  
    number_sections: false
    theme: flatly
    highlight: tango  
    toc_float:
      collapsed: true
      smooth_scroll: true
runtime: shiny
---

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(broom, wesanderson, latex2exp, leaflet, ggplot2, ggthemes, viridis, dplyr, magrittr, knitr, parallel, rddtools, readxl, emoGG, dslabs, gapminder, extrafont, Ecdat, wooldridge, tidyverse, janitor, kableExtra, gridExtra, estimatr, furrr, here, broom)# Define pink color
red_pink <- "#e64173"
turquoise <- "#20B2AA"
orange <- "#FFA500"
red <- "#fb6107"
blue <- "#3b3b9a"
green <- "#8bb174"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
slate <- "#314f4f"
# Dark slate grey: #314f4f
# Notes directory
# dir_slides <- "~/Dropbox/Courses/"
# Knitr options
# opts_chunk$set(
#   comment = "#>",
#   fig.align = "center",
#   fig.height = 4,
#   fig.width = 6,
#   # dpi = 300,
#   # cache = T,
#   warning = F,
#   message = F
# )
# A blank theme for ggplot
theme_empty <- theme_bw() + theme(
  line = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  axis.text = element_blank(),
  plot.title = element_blank(),
  axis.title = element_blank(),
  plot.margin = structure(c(0, 0, -1, -1), unit = "lines", valid.unit = 3L, class = "unit"),
  legend.position = "none"
)
theme_simple <- theme_bw() + theme(
  line = element_blank(),
  panel.grid = element_blank(),
  rect = element_blank(),
  axis.text.x = element_text(size = 10),
  axis.text.y = element_text(size = 10),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  axis.title.x = element_text(angle = 0, vjust = 0.5),
  axis.title.y = element_text(angle = 90, vjust = 0.5),
  legend.position = "none",
  axis.line = element_line(color="black", size = .5)
)
```

First, simulate a DGP in which there is some treatment to be evaluated. Do this in a way that allows you to control whether it falls randomly on individuals or systematically with some observable characteristic. 

Second, consider how one might display the existing variation in treatment associated with individual characteristics. A nice visualization---something you could display on your webpage when you're in the job market. 

```{r}
sim.dgp = function(n = 600) {
  tibble(
    i = 1:n,
    age = sample(c(24:80), n, replace = TRUE),
      age_bins = cut(age, breaks = c(24, 30, 45, 65, 81), right = FALSE),
    gender = sample(c('Male', 'Female'), n, replace = TRUE),
    female = ifelse(gender == 'Female', 1, 0),
    education = sample(c('High School', 'Some college', 'College', 'Advanced degree'), n, replace = TRUE, prob = c(0.40, 0.25, 0.20, 0.15)),
    
    college = ifelse(education == 'College' | education == 'Advanced degree', 1, 0),
    ## Introduce some age-based selection, different by gender...
    treated_num = ifelse(rnorm(n, sd = 1) + ifelse(female == 1, -age/45, age/60) < 0, 1, 0),
    treated = ifelse(treated_num == 1, 'Treated', 'Control'),
    y = 1 + female + college + treated_num + rnorm(n, sd = 1)
  )
}
```

```{r}
dgp = sim.dgp() %>% dplyr::select(c('age_bins', 'gender', 'education', 'y', 'treated'))
varSelectInput('Y_Axis', label = 'Select variable', data = dgp)
```

```{r}
checkboxInput("facetgender", "Facet wrap by gender?", FALSE)
```


```{r}
renderPlot({
g0 = ggplot(dgp, aes(x = treated, y = !!input$Y_Axis)) +
  theme_bw() + theme(
    line = element_blank(),
    panel.grid = element_blank(),
    rect = element_blank(),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.ticks = element_blank(),
    plot.title = element_blank(),
    axis.title.x = element_text(angle = 0, vjust = 0.5),
    axis.title.y = element_text(angle = 90, vjust = 0.5),
    legend.position = "right",
    legend.text.align = 0,
    axis.line = element_line(color = 'black', size = 0.5)
  ) +
  geom_count(aes(group = !!input$Y_Axis), color = red_pink) +
  scale_size_area(max_size = 12) +
  labs(x = " ",
       caption = paste0('Proportional breakdown of treatment status within selected variable, from simulated sample of....'),
       size = 'Proportion') +
  scale_color_manual(values = wes_palette(n = 3, name = 'Moonrise2'))

if (input$facetgender) {
  g0 + facet_grid(. ~ gender)
} else {
  g0
}
  
})
```

